Racial and Ethnic Differences in ETI: Outcomes

### 2024-11-14 Kamyron Jordan/Racial and Ethnic Differences in ETI/racial_and_ethnic_differences_in_eti_outcomes.qmd
just need contrasts for Kam and Scott

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
```

@@

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
cm <- chngptm(
  formula.1 = gli_fev1_ppred_rn ~ age_group + Insurance + (1 | eDWID),
  formula.2 = ~Years, family = "gaussian", data = encounter,
  type = "segmented", est.method = "grid", var.type = "none"
)

t <- data.frame(encounter)
levels(t$Race) <- c("White", "Black", "Hispanic", "Other")
o <- lme(gli_fev1_ppred_rn ~ Years + Race + age_group + Insurance,
  random = ~ 1 | eDWID, data = t, na.action = na.omit
)
os <- segmented(o, ~Years,
  data = t,
  fixed.parms = c(G0 = 0),
  random = list(eDWID = pdDiag(~ 1 + Years + U))
)

emt <- emtrends(change_mod, ~ Years + slope_change_years | Race, "Years",
  at = list(Years = c(-3, 2))
)
# EMMs
emm1 <- emmeans(change_mod,
  spec = ~ Years | Race,
  at = list(Years = c(-3, -2, -1, 0)),
  pbkrtest.limit = Inf
)
emm2 <- emmeans(change_mod,
  spec = ~ Years + slope_change_years | Race,
  at = list(Years = c(0, 1, 2), slope_change_years = c(0, 1, 2)),
  pbkrtest.limit = Inf
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
pop_dat <- mod_df %>% select(Years:Insurance)
pop_dat$pred <- predict(change_mod, re.form = NA, newdata = pop_dat)
pop_dat <- pop_dat %>% distinct()
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(data = pop_dat, aes(y = pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FEV1_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = GLI_FVC_pct_predicted, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
mod_df <- change_mod@frame
mod_df$ind_pred <- predict(change_mod)
mod_df$pop_pred <- predict(change_mod, re.form = NA)
ggplot(mod_df, aes(x = Years)) +
  geom_point(aes(y = gli_fev1fvc_ppred_rn, group = eDWID),
    alpha = 0.01, show.legend = F, shape = "."
  ) +
  geom_line(aes(y = ind_pred, group = eDWID),
    alpha = 0.01, show.legend = F
  ) +
  geom_line(aes(y = pop_pred, group = Race, colour = Race)) +
  xlab("Years from ETI Start") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression()
```

>>>>

```{r data import}
# source("/home/timvigers/GitHub/CF/Kamyron Jordan/Racial and Ethnic Differences in ETI/create_outcomes_dataset.R")
load("./Kamyron Jordan/Racial and Ethnic Differences in ETI/Data_Cleaned/outcomes_dataset.RData")
# Change age group reference level
encounter$age_group <- relevel(encounter$age_group, ref = "[18,30)")
```

@@

# Outcome distributions

### GLI 2022 equations

```{r}
ggplot(encounter, aes(x = gli_fev1_ppred_rn, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
ggplot(encounter, aes(x = gli_fev1_ppred_rn, fill = Race)) +
  geom_density(alpha = 0.2)
```

### GLI 2012 equations

```{r}
ggplot(encounter, aes(x = GLI_FEV1_pct_predicted, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
ggplot(encounter, aes(x = GLI_FEV1_pct_predicted, fill = Race)) +
  geom_density(alpha = 0.2)
```

### GLI 2022 equations

```{r}
ggplot(encounter, aes(x = gli_fvc_ppred_rn, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
ggplot(encounter, aes(x = gli_fvc_ppred_rn, fill = Race)) +
  geom_density(alpha = 0.2)
```

### GLI 2012 equations

```{r}
ggplot(encounter, aes(x = GLI_FVC_pct_predicted, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
ggplot(encounter, aes(x = GLI_FVC_pct_predicted, fill = Race)) +
  geom_density(alpha = 0.2)
```

### GLI 2022 equations

```{r}
ggplot(encounter, aes(
  x = gli_fev1fvc_ppred_rn, fill = Race,
  color = Race
)) +
  geom_density() +
  facet_wrap(~Race)
ggplot(encounter, aes(x = gli_fev1fvc_ppred_rn, fill = Race)) +
  geom_density(alpha = 0.2)
```

### GLI 2012 equations

```{r}
ggplot(encounter, aes(x = GLI_FEV1FVC_pct_predicted, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
ggplot(encounter, aes(x = GLI_FEV1FVC_pct_predicted, fill = Race)) +
  geom_density(alpha = 0.2)
```

## BMI %ile (age < 20 at ETI start)

```{r}
encounter %>%
  filter(age_eti_start < 20) %>%
  ggplot(aes(x = bmipercentile, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
encounter %>%
  filter(age_eti_start < 20) %>%
  ggplot(aes(x = bmipercentile, fill = Race)) +
  geom_density(alpha = 0.2)
```

## BMI raw value (age >= 20 at ETI start)

```{r}
encounter %>%
  filter(age_eti_start >= 20) %>%
  ggplot(aes(x = bmivalue, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
encounter %>%
  filter(age_eti_start >= 20) %>%
  ggplot(aes(x = bmivalue, fill = Race)) +
  geom_density(alpha = 0.2)
```

## Height %ile (age < 20 at ETI start)

```{r}
encounter %>%
  filter(age_eti_start < 20) %>%
  ggplot(aes(x = heightpercentile, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
encounter %>%
  filter(age_eti_start < 20) %>%
  ggplot(aes(x = heightpercentile, fill = Race)) +
  geom_density(alpha = 0.2)
```

## Height raw value (age >= 20 at ETI start)

```{r}
encounter %>%
  filter(age_eti_start >= 20) %>%
  ggplot(aes(x = height, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
encounter %>%
  filter(age_eti_start >= 20) %>%
  ggplot(aes(x = height, fill = Race)) +
  geom_density(alpha = 0.2)
```

## Weight %ile (age < 20 at ETI start)

```{r}
encounter %>%
  filter(age_eti_start < 20) %>%
  ggplot(aes(x = weightpercentile, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
encounter %>%
  filter(age_eti_start < 20) %>%
  ggplot(aes(x = weightpercentile, fill = Race)) +
  geom_density(alpha = 0.2)
```

## Weight raw value (age >= 20 at ETI start)

```{r}
encounter %>%
  filter(age_eti_start >= 20) %>%
  ggplot(aes(x = weight, fill = Race, color = Race)) +
  geom_density() +
  facet_wrap(~Race)
encounter %>%
  filter(age_eti_start >= 20) %>%
  ggplot(aes(x = weight, fill = Race)) +
  geom_density(alpha = 0.2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
pop_dat <- data.frame(expand.grid(
  Years = -3:2, Race = unique(encounter$Race), age_group = "[18,30)",
  Insurance = "Private"
))
pop_dat$slope_change_years <- (pop_dat$Years > 0) * pop_dat$Years
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FEV1_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FEV1_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FEV1_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race + slope_change_years * Race +
    Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[-Inf,18)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  GLI_FVC_pct_predicted ~ Years * Race +
    slope_change_years * Race + Insurance + (1 | eDWID),
  data = encounter[encounter$age_eti_group == "[18, Inf)", ]
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("GLI_FVC_pct_predicted") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(GLI_FVC_pct_predicted = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```

@@

```{r}
# Fit mixed model with change point
change_mod <- lmer(
  gli_fev1fvc_ppred_rn ~ Years * Race +
    slope_change_years * Race + age_group + Insurance + (1 | eDWID),
  data = encounter
)
# Model plot
pred <- predict(change_mod, re.form = NA, newdata = pop_dat, se.fit = T)
pop_dat$pred <- pred$fit
pop_dat$ll <- pop_dat$pred - (1.96 * pred$se.fit)
pop_dat$ul <- pop_dat$pred + (1.96 * pred$se.fit)
ggplot(pop_dat, aes(x = Years, y = pred, group = Race, colour = Race)) +
  geom_point() +
  geom_errorbar(aes(ymin = ll, ymax = ul), width = 0.1) +
  geom_line() +
  xlab("Years from ETI Start") +
  ylab("gli_fev1fvc_ppred_rn") +
  theme_classic()
# Model summary
change_mod %>% gtsummary::tbl_regression(intercept = T)
# Table of predicted values
pop_dat %>%
  select(-slope_change_years) %>%
  rename(gli_fev1fvc_ppred_rn = pred, ci.ll = ll, ci.ul = ul) %>%
  gt() %>%
  fmt_number(decimals = 2)
```
